---
title: "Lab 04 - La Quinta is Spanish for next to Denny's"
author: "Cailey Fay"
date: "10.13.25"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1
Each row represents a Denny's location. The variables are the address, city, state, zip code, longitude, and lattitude. There are  variables. 

### Exercise 2
The La Quinta dataset is The rows represent the motel, and the variables are the address, city, state, zip, longitude, and latitude. 

### Exercise 3
There are dennys outside of the US - Canada, Puerto Rico, UK. Same thing with La Quinta - they are in Canada, Mexico, Turkey, etc.

### Exercise 4
I would look at longitude and latitude and do a filter so that we isolate to just the rectangle encompassing all of the US minus Hawaii and Alaska and then I'd probably manually add in the Alaska / Hawaii because how many could there really be? 

### Exercise 5
There are no Denny's outside of the US in this data file
```{r excercise5}
dennys %>%
  filter(!(state %in% states$abbreviation))
```


### Exercise 6
```{r exercise6}
dn <- dennys %>%
    mutate(country="United States")
```


### Exercise 7
There are 14 laquintas outside of the US. They belong to the following countries: Mexico x 10, colombia, Canada x2, Honduras. Abbreviations are AG, QR, CH, NL, ANT, ON, VE, PU, SL, FM, and BC.
```{r exercise7}
internationallaquinta <- laquinta %>%
  filter(!(state %in% states$abbreviation))

```


### Exercise 8
``` {r exercise8}
newlaquinta <- laquinta %>%
  mutate(country = case_when(
  state %in% state.abb ~ "United States",
  state %in% c("ON", "BC") ~ "Canada",
  state == "ANT" ~ "Colombia",
  state %in% c("AG", "QR", "CH", "NL", "PU","SL", "FM") ~ "Mexico",
  state == "FM"~"Honduras"
  ))
  
lq <- newlaquinta %>%
  filter(country == "United States")
```

### Exercise 9 
California and Texas have the most, at 403 and 200, respectively.  
```{r exercise9}
dn %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  arrange(desc(n))
```


### Exercise 10
DC has the most Denny's, with 29 per 1000 square miles. This is a little misleading, though, because of how tiny DC is. DC only has 2 Dennys, and RI comes in second place with 5 Dennys. The third place for number of denny's per 1000 square miles is California, which has 403 Dennys. 
```{r}
dn_state <- dn %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  arrange(desc(n)) 

dn_state <- dn_state %>%
  mutate(dn_count = n / area)

dn_state <- dn_state %>%
  mutate(dn_per1kmiles = dn_count * 1000)

dn_state %>%
  arrange(desc(dn_per1kmiles))
```
Rhode island has the most La Quinta's per 1000k miles, followed by Florida and Connecticut.  
```{r}
lq_state <- lq %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  arrange(desc(n)) 

lq_state <- lq_state %>%
  mutate(lq_count = n / area)

lq_state <- lq_state %>%
  mutate(lq_per1kmiles = lq_count * 1000) 

lq_state %>%
  arrange(desc(lq_per1kmiles))
```
Mapping Denny's and La Quintas in the US:
```{r exercise10}
dn <- dn %>%
  mutate(establishment = "Denny's")
lq <- lq %>%
  mutate(establishment = "La Quinta")
dn_lq <- bind_rows(dn, lq)
# made the new combined variable successfully 
ggplot(dn_lq, mapping = aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point(alpha = .85) +
  labs(title = "Denny's and La Quinta Locations Across the United States",
       x = "Longitude",
       y = "Latitude",
       color = "Establishment")
```

### Exercise 11
```{r exercise11}
NConly <- dn_lq %>%
  filter(state == "NC")
ggplot(NConly, mapping = aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point() +
    labs(title = "Denny's and La Quinta Locations In North Carolina",
       x = "Longitude",
       y = "Latitude",
       color = "Establishment")
```

### Exercise 12
The joke most certainly holds for Texas, but only sort of for NC. 
```{r exercise12}
TXonly <- dn_lq %>%
  filter(state == "TX")
ggplot(TXonly, mapping = aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point(alpha = .75) +
    labs(title = "Denny's and La Quinta Locations In Texas",
       x = "Longitude",
       y = "Latitude",
       color = "Establishment")

```
